---
title: 'US vs World: China/Hunan Virus 2019'
author: "David J Jackson"
date: "Apr. 17, 2020 01:35PM EDT"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DT)
library(pracma)
library(plotly)
library(scales)
library(prophet)
library(readxl)
library(httr)
library(lubridate)
library(RSQLite)
```
```{r, echo=FALSE}
rm(list=ls())
#these libraries are necessary
#create the URL where the dataset is stored with automatic updates every day

url <- paste("https://www.ECDC.europa.eu/sites/default/files/documents/COVID-19-geographic-disbtribution-worldwide-",format(Sys.time(), "%Y-%m-%d"), ".xlsx", sep = "")

#download the dataset from the website to a local temporary file
GET(url, authenticate(":", ":", type="ntlm"), write_disk(tf <- tempfile(fileext = ".xlsx")))
#read the Dataset sheet into “R”
ECDC <- read_excel(tf)
colnames(ECDC) <- c("Reported","Day","Month","Year","Cases","Deaths","Countries","GeoId","CountryCode","Population")
```

#### Data Date Range
```{r,echo=FALSE}
min(ECDC$Reported)
max(ECDC$Reported)
```
```{r,echo=FALSE}
# ECDC$DateRep <- lubridate::ymd(ECDC$Reported)
```
```{r,echo=FALSE}

# ECDC$DateRep <- as.Date(ECDC$DateRep)
# ECDC$DCount <- ifelse(ECDC$Deaths ==0,"NO","YES")
ECDC$Countries <-ifelse(ECDC$Countries=="United_States_of_America","USA",ECDC$Countries)
ECDC$Countries <-ifelse(ECDC$Countries=="Cases_on_an_international_conveyance_Japan","Japan",ECDC$Countries)
```
```{r,echo=FALSE}
# Calculate Summary Cases and Deaths(World)
# World
# ECDC Summary by Year
ECDC_Year <- ECDC %>% group_by(Yearly =floor_date( Reported,"year")) %>% summarise(
                                                        tc = sum(Cases),
                                                        td = sum(Deaths),                                                                perdeaths = td/tc,
                                                        Count = n())

# Summary by Month
ECDC_monthly <- ECDC %>% group_by(Monthly =floor_date( Reported,"month")) %>%                                                              summarise(
                                                        tc = sum(Cases),
                                                        td = sum(Deaths),                                                                perdeaths = td/tc,
                                                        Count = n())

## Weekly Summary
ECDC_weekly <- ECDC %>% group_by(Weekly =floor_date( Reported,"week")) %>% summarise(
                                                        tc = sum(Cases),
                                                        td = sum(Deaths),                                                                 perdeaths = td/tc,
                                                        Count = n())
```

```{r, echo=FALSE}
# By Country
# Weekly
Countries_yearly <- ECDC %>% group_by(Countries,Yearly =floor_date( Reported,"year")) %>% summarise(
                                                        tc = sum(Cases),
                                                        td = sum(Deaths),                                                                 perdeaths = td/tc,
                                                              Count = n())

# Monthly
# Monthly
Countries_monthly <- ECDC %>% group_by(Countries,Monthly =floor_date( Reported,"month")) %>% summarise(
                                                        tc = sum(Cases),
                                                        td = sum(Deaths),                                                                 perdeaths = td/tc,
                                                              Count = n())

# Weekly
Countries_weekly <- ECDC %>% group_by(Countries,Weekly =floor_date( Reported,"week")) %>% summarise(
                                                        tc = sum(Cases),
                                                        td = sum(Deaths),                                                                 perdeaths = td/tc,
                                                              Count = n())
# World minus US
ECDC_minus_us <- Countries_weekly %>% filter(Countries != "USA" )

# USA Only
US_yearly <- Countries_yearly %>% filter(Countries == "USA" )
US_monthly <- Countries_monthly %>% filter(Countries == "USA" )
US_weekly <- Countries_weekly %>% filter(Countries == "USA" )
```

### The World: Summary by Week for Cases,Deaths, and Percent of Deaths

```{r, echo=FALSE}
ECDC_weekly %>% plot_ly() %>% add_lines(x=~Weekly,y=~tc) %>% 
        layout(title="The World: Total Cases By Week")


ECDC_weekly %>% plot_ly() %>% add_lines(x=~Weekly,y=~td) %>% 
        layout(title="The World: Total Deaths By Week")

ECDC_weekly %>% plot_ly() %>% add_lines(x=~Weekly,y=~perdeaths) %>% 
        layout(title="The World: Percent of Cases Ending in Death by Week") 
      
```

### US Summary by Week of Cases,Deaths and Percent of Deaths

```{r,echo=FALSE}

US_weekly %>% plot_ly() %>% add_lines(x=~Weekly,y=~tc) %>% 
        layout(title="USA: Total Cases By Week")
US_weekly %>% plot_ly() %>% add_lines(x=~Weekly,y=~td) %>% 
        layout(title="USA: Total Deaths By Week")
US_weekly %>% plot_ly() %>% add_lines(x=~Weekly,y=~perdeaths) %>% 
        layout(title="USA Percent of Cases Ending in Death by Week")
```

## US vs World: Percent of Death by Week

```{r,echo=FALSE}
plot_ly() %>% add_lines(data=US_weekly,x=~Weekly,y=~perdeaths,name="USA") %>% 
        add_lines(data=ECDC_weekly,x=~Weekly,y=~perdeaths,name="World") %>%
        layout(title="USA/World Percent of Cases Ending in Death by Week")
```

### The World minus US: Summary by Week of Cases,Deaths, and Percent of Deaths

```{r, echo=FALSE}

ECDC$Reported <- as.character(ECDC$Reported)
db <- dbConnect(SQLite(), dbname="../COVIDDB//CORVID.sqlite3")
```
```{r,echo=FALSE}
dbWriteTable(db, "ECDC", ECDC,overwrite=TRUE)
```
```{r, echo=FALSE}
# GEOID <- ECDC %>% group_by(Countries,CountryCode,Population) %>% 
#         summarise(Count=n())
# dbWriteTable(db, "GEOID", GEOID,overwrite=TRUE)
dbListTables(db)
dbDisconnect(db)
```

### Daly Totals form Cases,Deaths and Death Rate

```{r, echo=FALSE}
df <-ECDC %>% group_by(Reported) %>% summarise(
                                          TotalCases=sum(Cases),
                                         TotalDeaths = sum(Deaths),
                                         DeathRate = TotalDeaths/TotalCases)
df %>% plot_ly() %>% add_bars(x=~Reported,y=~TotalCases) %>% layout(title="Total Confrimed Cases by Day")

df %>% plot_ly() %>% add_bars(x=~Reported,y=~TotalDeaths) %>% layout(title="Total Confrimed Deaths by Day")

df %>% plot_ly() %>% add_bars(x=~Reported,y=~DeathRate) %>% layout(title="Total Confrimed Deaths by Day")
```

### US Daily Reports for Cases,Deaths and Death Rate

```{r,echo=FALSE}

df.us <-ECDC %>% filter(GeoId =="US") %>% group_by(Reported) %>% summarise(
                                          TotalCases=sum(Cases),
                                         TotalDeaths = sum(Deaths),
                                         DeathRate = TotalDeaths/TotalCases)

df.us %>% plot_ly() %>% add_bars(x=~Reported,y=~TotalCases) %>% layout(title="US Total Confrimed Cases by Day")

df.us %>% plot_ly() %>% add_bars(x=~Reported,y=~TotalDeaths) %>% layout(title="US Total Confrimed Deaths by Day")

df.us %>% plot_ly() %>% add_bars(x=~Reported,y=~DeathRate) %>% layout(title="US Death Rate (as Percentage) by Day")
